\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2022-01-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R1.6. Comments by reviewer}:\\
{\em The simulation time is too short to convince the readers. Only 1 second is considered here. Note that the errors can be quickly accumulated in the iteration.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.7. Comments by reviewer}:\\
{\em It is suggested to count the total computing time, not on each step.}

{\nib{ \blue While the depiction of the per-step computational times in Figs. 8 and 9 in the revised manuscript are somewhat unconventional, we explain why we choose to plot that way. \ul{To aid clarity, in the revised manuscript, we have added an excerpt of the following at the end of Sec. V-A.}

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. \ul{To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time}. On the other hand, theoretical consistency (see the unnumbered equation after (26), before Sec. IV) requires that $h$ itself should be small. Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear \emph{a priori} if the computational time needed to do so can be much faster than $h$. If not, then the method would be statistically consistent but practically not useful. The Figs. 8 and 9 in the revised manuscript show that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.}}

### 2022-01-10 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

(2) \textbf{Regarding randomness in active and reactive power:} {\red{TBD}}

(3) \textbf{Regarding the diffusion term:} The reason why the diffusion term is present in the RHS of (1b) but not in the RHS of (1a) is because (1) is the state space representation of the second order Langevin dynamics (6). For the Langevin dynamics (6), its corresponding It\^{o} SDE \ul{has to be} (1). This is simply re-writing of a second order differential equation as an equivalent system of first order differential equations. There is no extra assumption here. \ul{Indeed, it is mathematically incorrect to add a noise term in the RHS of (1a) while claiming consistency with (6)}.

\ul{Notice that the absence of additive noise in the RHS of (1a) does \emph{not} imply that the rotor angles are deterministic trajectories}. The reason why the rotor angles are stochastic processes is because the process noise in (1b) affects the rotor angle sample paths via dynamical nonlinear coupling with the angular velocities, i.e., these are second order stochastic effects. \ul{In short, the noise in rotor angles is \emph{not} ignored; in fact the rotor angles are nonlinearly colored (i.e., non-white) noisy sample paths}.

This reviewer may also have \ul{mathematical misconception} in thinking that a model that has additive noise in the RHS of \emph{all} SDE components is somehow more general than the case where the additive noise terms are absent from some but not all components of an SDE RHS. \ul{In reality, the complexity is in the other way around}. The SDEs of the form (1) are the so-called ``degenerate It\^{o} SDEs" which are much more difficult to analyze than the non-degenerate versions (noise in the RHS of all SDE components). For instance, this degeneracy makes the joint PDF evolution equation (10) a \emph{kinetic} Fokker-Planck PDE, which is different in structure than the standard Fokker-Planck a.k.a. Kolmogorov's forward PDE. \ul{The gradient flow structure and the associated measure-valued proximal recursions for the kinetic Fokker-Planck PDE are less understood than the same for the standard Fokker-Planck PDE. This is exactly why the methodological contribution of this paper is important, even if one completely ignores the algorithmic contribution}.

@@

\noindent
{\bf R1.4. Comments by reviewer}:\\
{\em There is no comparison of the proposed method compared with other state-of-the-art methods. Therefore, it is hard for the readers to judge the importance of this article.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.5. Comments by reviewer}:\\
{\em Contributions are over-claimed in this article:\\
(1) The scalability of the proposed method cannot convince the readers since you are using a too simple model and ignored many important random dimensions from the loads.\\
(2) The proposed method claims to be nonparametric, however, all the inputs are assumed to be either wiener process or uniform distribution, which are all parametric distributions.\\
(3) If the ``nonparametric" is only referred to the output PDF, I would not think of it as a major contribution since there are a lot of existing techniques that can depict nonparametric PDF.}

(1) \textbf{Regarding simple model and ignoring the randomness in loads}: Please refer to our response to \textbf{R1.3, sub-comment (1)} where we explained why this model is relevant for this particular problem, and why this is not limiting to our methodological contribution. In response to \textbf{R1.3, last unnumbered sub-comment in bold text: ``Regarding OU process in load"}, we explained that we are not ignoring randomness from loads.

(2) \textbf{Regarding nonparametric}: \ul{We are \emph{not} making any parametric assumptions}. Our usage of the word ``nonparametric" is in the standard sense used in mathematical statistics literature. That the additive noise process is Wiener does not invite any loss of generality since if it was colored, we could always re-cast the same in our SDE form via a linear map. \ul{Furthermore, we have zero assumptions on the initial joint state PDF $\rho_{0}$ and on the joint PDF of the parameters appearing in the model. As a matter of fact, we don't even need to know the functional form of these joint PDFs. For instance, our framework remains unchanged if we are given the initial joint state and parameter distributional data as numerical table containing samples from these joints \emph{as well as} the joint PDF values evaluated at those samples}. The latter might be coming from some measurement-based sensor fusion/Bayesian algorithm. The specific product von Mises and uniform PDFs in Sec. V-A and V-B are only used for generating data to perform our numerical example simulations. \ul{This again underscores how a reader jumping straight to Sec. V without contemplative reading and understanding of the earlier Sections will make prematurely incorrect claims}.

(3) \textbf{Regarding depict nonparametric PDF}: \ul{This paper is not about nonparametric \emph{depiction} but about nonparametric \emph{computation} of the transient joint state PDFs}. For example, Monte Carlo followed by function approximation via Kernel density estimates and its variants, or other function approximants (e.g., RBF, Gaussian mixture, neural networks) are all parametric methods. We enable computing the value of joint state PDF at a time-varying sample location \emph{online (not as offline post-processing)} without assuming any functional representation of the transient joint which the computed values correspond to. Notice that we are doing this joint computation without gridding the state space, which would be impossible had we followed the traditional scientific computing route: finite element/volume methods, tensor decomposition techniques etc. \ul{We believe the proposed computational framework is novel, highly non-trivial and is specifically useful in this power system uncertainty propagation context}.
}}

### 2022-01-10 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R1.5. Comments by reviewer}:\\
{\em Contributions are over-claimed in this article:\\
(1) The scalability of the proposed method cannot convince the readers since you are using a too simple model and ignored many important random dimensions from the loads.\\
(2) The proposed method claims to be nonparametric, however, all the inputs are assumed to be either wiener process or uniform distribution, which are all parametric distributions.\\
(3) If the ``nonparametric" is only referred to the output PDF, I would not think of it as a major contribution since there are a lot of existing techniques that can depict nonparametric PDF.}

{\nib{ \blue \ul{Our responses to the sub-comments below}.

(1) \textbf{Regarding simple model and ignoring the randomness in loads}: Please refer to our response to \textbf{R1.3, sub-comment (1)} where we explained why this model is relevant for this particular problem, and why this is not limiting to our methodological contribution. In response to \textbf{R1.3, last unnumbered sub-comment in bold text: ``Regarding OU process in load"}, we explained that we are not ignoring randomness from loads.

(2) \textbf{Regarding nonparametric}: \ul{We are \emph{not} making any parametric assumptions}. Our usage of the word ``nonparametric" is standard in mathematical statistics literature. That the additive noise process is Wiener does not invite any loss of generality since if it was colored, we could always re-cast the same in our SDE form via a linear map. \ul{Furthermore, we have zero assumptions on the initial joint state PDF $\rho_{0}$ and on the joint PDF of the parameters appearing in the model. As a matter of fact, we don't even need to know the functional form of these joint PDFs. For instance, our framework remains unchanged if we are given the initial joint state and parameter distributional data as numerical table containing samples from these joints \emph{as well as} the joint PDF values evaluated at those samples}. The latter might be coming from some measurement-based sensor fusion/Bayesian algorithm. The specific product von Mises and uniform PDFs in Sec. V-A and V-B are only used for generating example data to perform our illustrative numerical simulations. \ul{This again underscores how a reader jumping straight to Sec. V without contemplative reading and understanding of the earlier Sections may make prematurely incorrect claims}.

### 2022-02-19 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\section*{\large \bf Response to Reviewer \#1:}

\noindent
{\bf R1.1. Comment by reviewer}:\\
{\em The original contribution of the article is not clear.  Indeed, what is original new from this article, not the summary from existing techniques?}

{\nib {\blue The original contribution of this article is to design an algorithm for \emph{online nonparametric} computation of the joint state PDFs subject to stochastic uncertainties (initial conditions, parameters, process noise and combinations thereof) in power system dynamics.

The qualifier ``online" means that it is not a variant of Monte Carlo where samples are first propagated \ul{and then} a function approximation is performed offline to estimate the joint state PDF at a snapshot. The qualifier ``nonparametric" means that neither the statistics nor the dynamics are approximated \emph{a priori}. This is novel because the existing power systems literature makes at least one of the aforesaid approximations. The state-of-the-art in power system literature is discussed in Sec I.1. More importantly, Sec. I.2 details why the state-of-the-art is what it is. \ul{We specifically explain the contribution of this article in Sec. I.3 titled ``Contributions of this paper".}
}}

@@

(2) \textbf{Regarding randomness in active and reactive power:} {\red{TBD}}

(3) \textbf{Regarding the diffusion term:} The reason why the diffusion term is present in the RHS of (1b) but not in the RHS of (1a) is because (1) is the state space representation of the second order Langevin dynamics (6). For the Langevin dynamics (6), its corresponding It\^{o} SDE \ul{has to be} (1). This is simply re-writing of a second order differential equation as an equivalent system of first order differential equations. There is no extra assumption here. \ul{Indeed, it is mathematically incorrect to add a noise term in the RHS of (1a) while claiming consistency with (6)}.

\ul{Notice that the absence of additive noise in the RHS of (1a) does \emph{not} imply that the rotor angles are deterministic trajectories}. The reason why the rotor angles are stochastic processes is because the process noise in (1b) affects the rotor angle sample paths via dynamical nonlinear coupling with the angular velocities, i.e., these are second order stochastic effects. \ul{In short, the noise in rotor angles is \emph{not} ignored; in fact the rotor angles are nonlinearly colored (i.e., non-white) noisy sample paths}.

This reviewer may also have misunderstood in thinking that a model that has additive noise in the RHS of \emph{all} SDE components is somehow more general than the case where the additive noise terms are absent from some but not all components of an SDE RHS. \ul{In reality, the complexity is in the other way around}. The SDEs of the form (1) are the so-called ``degenerate It\^{o} SDEs" which are much more difficult to analyze than the non-degenerate versions (noise in the RHS of all SDE components). For instance, this degeneracy makes the joint PDF evolution equation (10) a \emph{kinetic} Fokker-Planck PDE, which is different in structure than the standard Fokker-Planck a.k.a. Kolmogorov's forward PDE. \ul{The gradient flow structure and the associated measure-valued proximal recursions for the kinetic Fokker-Planck PDE are less understood than the same for the standard Fokker-Planck PDE. This is exactly why the methodological contribution of this paper is important, even if one completely ignores the algorithmic contribution}.

\textbf{Regarding the reference [R1]:} \ul{In the revised manuscript, we have included the ref. [R1] suggested by the reviewer with inline citation in Sec. I.1, toward the end of second paragraph. Please note that the topic of [R1] is different from the topic of our paper. We are interested in how to perform online nonparametric computation of transient joint state PDFs}. Our computation is different from Monte Carlo.

\textbf{Regarding OU process in load:} \ul{The reviewer might have missed the following key point of our paper}. The Kron reduced second order stochastic model implies that in the unreduced sense, the load dynamics was first order noisy Kuramoto, in addition to the synchronous machine dynamics being second order noisy Kuramoto. Therefore, we do have stochastic load dynamics, which is in fact more realistic than OU process model. While OU process assumes linear drift with constant diffusion, whereas our setting considers the load SDEs having drifts with trigonometric nonlinearity and constant diffusion.

@@

\noindent
{\bf R1.5. Comments by reviewer}:\\
{\em Contributions are over-claimed in this article:\\
(1) The scalability of the proposed method cannot convince the readers since you are using a too simple model and ignored many important random dimensions from the loads.\\
(2) The proposed method claims to be nonparametric, however, all the inputs are assumed to be either wiener process or uniform distribution, which are all parametric distributions.\\
(3) If the ``nonparametric" is only referred to the output PDF, I would not think of it as a major contribution since there are a lot of existing techniques that can depict nonparametric PDF.}

{\nib{ \blue \ul{Our responses to the sub-comments below}.

(1) \textbf{Regarding simple model and ignoring the randomness in loads}: Please refer to our response to \textbf{R1.3, sub-comment (1)} where we explained why this model is relevant for this particular problem, and why this is not limiting to our methodological contribution. In response to \textbf{R1.3, last unnumbered sub-comment in bold text: ``Regarding OU process in load"}, we explained that we are not ignoring randomness from loads.

(2) \textbf{Regarding nonparametric}: \ul{We are \emph{not} making any parametric assumptions}. Our usage of the word ``nonparametric" is standard as in mathematical statistics literature. That the additive noise process is Wiener does not invite any loss of generality since if it was colored, we could always re-cast the same in our SDE form via a linear map. \ul{Furthermore, we have zero assumptions on the initial joint state PDF $\rho_{0}$ and on the joint PDF of the parameters appearing in the model. As a matter of fact, we don't even need to know the functional form of these joint PDFs. For instance, our framework remains unchanged if we are given the initial joint state and parameter distributional data as numerical table containing samples from these joints \emph{as well as} the joint PDF values evaluated at those samples}. The latter might be coming from some measurement-based sensor fusion/Bayesian algorithm. The specific product von Mises and uniform PDFs in Sec. V-A and V-B are only used for generating example data to perform our illustrative numerical simulations. \ul{This again underscores how jumping straight to Sec. V without contemplative reading and understanding of the earlier Sections may lead to misleading conclusions}.

@@

(3) \textbf{Regarding depict nonparametric PDF}: \ul{This paper is not about nonparametric \emph{depiction} but about nonparametric \emph{computation} of the transient joint state PDFs}. For example, Monte Carlo followed by function approximation via Kernel density estimates and its variants, or other function approximants (e.g., RBF, Gaussian mixture, neural networks) are all parametric methods. We enable computing the value of joint state PDF at a time-varying sample location \emph{online (not as offline post-processing)} without assuming any functional representation of the transient joint which the computed values correspond to. Notice that we are doing this joint computation without gridding the state space, which would be impossible had we followed the traditional scientific computing route: finite element/volume methods, tensor decomposition techniques etc. \ul{We believe the proposed computational framework is novel, highly non-trivial and is specifically useful in this power system uncertainty propagation context}.
}}

\noindent
{\bf R1.6. Comments by reviewer}:\\
{\em The simulation time is too short to convince the readers. Only 1 second is considered here. Note that the errors can be quickly accumulated in the iteration.}

{\nib{ \blue Please note that the convergence guarantee given by equation (14) in the revised manuscript is pointwise in time, i.e., for each $k\in\mathbb{N}$. Unlike traditional numerical methods, there is no concept of {\red{TBD.}}}}

### 2022-02-19 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

\begin{abstract}
We present a proximal algorithm that performs a variational recursion on the space of joint probability measures to propagate the stochastic uncertainties in power system dynamics over high dimensional state space. The proposed algorithm takes advantage of the exact nonlinearity structures in the trajectory-level dynamics of the networked power systems, and is nonparametric. Lifting the dynamics to the space of probability measures allows us to design a scalable algorithm that obviates gridding the underlying high dimensional state space which is computationally prohibitive. The proximal recursion implements a generalized infinite dimensional gradient flow, and evolves probability weighted scattered point clouds. We clarify the theoretical nuances and algorithmic details specific for the power system nonlinearities, and provide illustrative numerical examples.
\end{abstract}

@@

%The study of power system dynamics with uncertainties is not new. However, 
\IEEEPARstart{S}{tochastic} variabilities in power grid have increased significantly in recent years both in the generation side (e.g., due to growing penetration of renewables) as well as in the load side (e.g., due to widespread adoption of plug-in electric vehicles). Several studies \cite{timko1983monte,nwankpa1992stochastic,odun2012structure,ghanavati2016identifying,apostolopoulou2016assessment} have reported that even small stochastic effects can significantly alter the assessment of transient stability, or the performance of automatic generation control. However, the lack of a rigorous yet scalable stochastic computational framework continues to impede \cite{schwalbe2015mathematical} our ability to perform transient analysis involving time varying joint probability density functions (PDFs) over the states of a large power system network. In this paper, we present a new algorithm to address this computational need.

@@

 As is well known, the Monte Carlo techniques are easy to apply but the computational cost scales exponentially with the number of dimensions, thus making it prohibitive for realistic power systems dynamic simulation. As an alternative, \emph{probabilistic small signal analysis} \cite{nwankpa1992stochastic,rueda2009assessment,huang2013quasi,dhople2013analysis} have appeared in the power systems literature, albeit at the expense of the additional assumption that the random perturbations remain ``small". \emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite-dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired. Furthermore, to cope with the stochasticity, these techniques require simulating a higher-dimensional system than the dimension of the physical state space, which further limits the scalability for nonlinear simulation. More recently, approximation methods such as \emph{stochastic averaging} \cite{ju2018analytical,ju2018stochastic} based on certain energy function \cite{pai1989energy,chang1995direct,sauerpai1998} have appeared. In \cite{maldonado2018uncertainty}, an algorithm for \emph{propagating first few statistical moments} was proposed. However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general. {\blue{We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}. These studies, however, simulate the sample paths and do not directly propagate the joint state PDFs.}}

In a different vein, \emph{deterministic bounded uncertainty models} for power flow simulations have been used \cite{dimitrovski2004boundary,hiskens2006sensitivity,chen2012method,althoff2014formal,geng2020reach} for set-valued analysis. These, however, require approximating the underlying nonlinear (deterministic) DAEs appearing in power system dynamic simulation, and thus lead to conservative analysis. For example, the method in \cite{althoff2014formal} requires converting the nonlinear DAEs to linear DAEs in such a way that guarantees set-valued over-approximation of the reachable sets. Likewise, the convex optimization-based bounded uncertainty propagation methods in \cite{choi2017propagating,choi2018propagating} require second order approximation of the power flow state variables as function of the uncertainties.

@@

Our technical approach juxtaposes with the existing power systems literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics. Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity over the manifold of time-varying joint state PDFs, thereby enabling a new proximal algorithm to compute the transient joint PDFs.

\subsubsection{Notation and organization}
%We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. 
The set of natural numbers, real numbers and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. We use the superscript $*$ to denote the complex conjugate, and the superscript $^{\top}$ to denote the matrix transpose. The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$. Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$. The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute magnitude, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

@@

\begin{figure}[t]
\centering
\includegraphics[width=.7\linewidth]{BlockDiagm.pdf}
\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$. The states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec. \ref{sec:ProxAlgo}.}}
\label{fig:BlockDiagm}
\vspace*{-0.2in}
\end{figure}

@@

For completeness, the algorithm \textproc{Prox} for updating $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ is outlined in Algorithm \ref{algo:KineticProx}. As shown in Fig. \ref{fig:BlockDiagm}, this algorithm, at a conceptual level, takes the pre and post update state samples 
\[\{\tilde{\bm{x}}_{k-1}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k-1}^{i},\bm{\eta}_{k-1}^{i}\right)\}_{i=1}^{N}, \; \{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)\}_{i=1}^{N},\]
and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs, and outputs the proximal updates $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$. For each $k\in\mathbb{N}$, the updated probability weighted point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are then brought back to the original state space via $\bm{\Psi}^{-1}$ as $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$, as explained earlier.

The Algorithm \ref{algo:KineticProx} also needs input parameters $h,\varepsilon,N,\delta,L$. Specifically, $h$ is the time-step size, $\varepsilon$ is an entropic regularization weight internal to the \textproc{Prox} algorithm, and $N$ is the number of samples. The parameters $\delta$ and $\ell_{\max}$ codify the numerical tolerance and maximum number of iterations, respectively, for the while loop in Algorithm \ref{algo:KineticProx}. Its convergence guarantees can be found in \cite[Sec. III.C]{caluya2019TAC}.

@@

\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and assuming to remain constant over the simulation time horizon. The transient stochastic simulations (i.e., with process noise) are then performed using the proposed framework.

>>>>

\begin{abstract}
We present a proximal algorithm that performs a variational recursion on the space of joint probability measures to propagate the stochastic uncertainties in power system dynamics over high dimensional state space. The proposed algorithm takes advantage of the exact nonlinearity structures in the trajectory-level dynamics of the networked power systems, and is nonparametric. Lifting the dynamics to the space of probability measures allows us to design a scalable algorithm that obviates gridding the underlying high dimensional state space which is computationally prohibitive. The proximal recursion implements a generalized infinite dimensional gradient flow, and evolves probability-weighted scattered point clouds. We clarify the theoretical nuances and algorithmic details specific to the power system nonlinearities, and provide illustrative numerical examples.
\end{abstract}

@@

%The study of power system dynamics with uncertainties is not new. However, 
\IEEEPARstart{S}{tochastic} variabilities in power grid have increased significantly in recent years both in the generation side (e.g., due to growing penetration of renewables) as well as in the load side (e.g., due to widespread adoption of plug-in electric vehicles). Several studies \cite{timko1983monte,nwankpa1992stochastic,odun2012structure,ghanavati2016identifying,apostolopoulou2016assessment} have reported that even small stochastic effects can significantly alter the assessment of transient stability, or the performance of automatic generation control. However, the lack of a rigorous yet scalable stochastic computational framework continues to impede \cite{schwalbe2015mathematical} our ability to perform transient analysis involving time-varying joint probability density functions (PDFs) over the states of a large power system network. In this paper, we present a new algorithm to address this computational need.

@@

 As is well known, the Monte Carlo techniques are easy to apply but the computational cost scales exponentially with the number of dimensions, thus making it prohibitive for realistic power systems dynamic simulation. As an alternative, \emph{probabilistic small signal analysis} \cite{nwankpa1992stochastic,rueda2009assessment,huang2013quasi,dhople2013analysis} have appeared in the power systems literature, albeit at the expense of the additional assumption that the random perturbations remain ``small". \emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired. Furthermore, to cope with the stochasticity, these techniques require simulating a higher-dimensional system than the dimension of the physical state space, which further limits the scalability for nonlinear simulation. More recently, approximation methods such as \emph{stochastic averaging} \cite{ju2018analytical,ju2018stochastic} based on certain energy function \cite{pai1989energy,chang1995direct,sauerpai1998} have appeared. In \cite{maldonado2018uncertainty}, an algorithm for \emph{propagating first few statistical moments} was proposed. However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general. {\blue{We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}. These studies, however, simulate the sample paths and do not directly propagate the joint state PDFs.}}

In a different vein, \emph{deterministic bounded uncertainty models} for power flow simulations have been used \cite{dimitrovski2004boundary,hiskens2006sensitivity,chen2012method,althoff2014formal,geng2020reach} for set-valued analysis. These, however, require approximating the underlying nonlinear (deterministic) DAEs appearing in power system dynamic simulation, and thus lead to conservative analysis. For example, the method in \cite{althoff2014formal} requires converting the nonlinear DAEs to linear DAEs in such a way that guarantees set-valued over-approximation of the reachable sets. Likewise, the convex optimization-based bounded uncertainty propagation methods in \cite{choi2017propagating,choi2018propagating} require second order approximation of the power flow state variables as a function of the uncertainties.

@@

Our technical approach juxtaposes with the existing power systems literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics. Instead of treating the exact nonlinearity as a bane, we exploit the geometry induced by the power system's dynamical nonlinearity over the manifold of time-varying joint state PDFs, thereby enabling a new proximal algorithm to compute the transient joint PDFs.

\subsubsection{Notation and organization}
%We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. 
The set of natural numbers, real numbers, and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. We use the superscript $*$ to denote the complex conjugate, and the superscript $^{\top}$ to denote the matrix transpose. The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$. Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$. The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute magnitude, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

@@

\begin{figure}[t]
\centering
\includegraphics[width=.7\linewidth]{BlockDiagm.pdf}
\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability-weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$. The states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec. \ref{sec:ProxAlgo}.}}
\label{fig:BlockDiagm}
\vspace*{-0.2in}
\end{figure}

@@

For completeness, the algorithm \textproc{Prox} for updating $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ is outlined in Algorithm \ref{algo:KineticProx}. As shown in Fig. \ref{fig:BlockDiagm}, this algorithm, at a conceptual level, takes the pre and post-update state samples 
\[\{\tilde{\bm{x}}_{k-1}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k-1}^{i},\bm{\eta}_{k-1}^{i}\right)\}_{i=1}^{N}, \; \{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)\}_{i=1}^{N},\]
and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs, and outputs the proximal updates $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$. For each $k\in\mathbb{N}$, the updated probability-weighted point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are then brought back to the original state space via $\bm{\Psi}^{-1}$ as $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$, as explained earlier.

Algorithm \ref{algo:KineticProx} also needs input parameters $h,\varepsilon,N,\delta,L$. Specifically, $h$ is the time-step size, $\varepsilon$ is an entropic regularization weight internal to the \textproc{Prox} algorithm, and $N$ is the number of samples. The parameters $\delta$ and $\ell_{\max}$ codify the numerical tolerance and maximum number of iterations, respectively, for the while loop in Algorithm \ref{algo:KineticProx}. Its convergence guarantees can be found in \cite[Sec. III.C]{caluya2019TAC}.

@@

\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and {\blue{are assumed}} to remain constant over the simulation time horizon. The transient stochastic simulations (i.e., with process noise) are then performed using the proposed framework.
